### 2013-07-06 PaperDraft.Rmd
more advances in the paper.

Learning Polynomial Regression with Genetic Algorithms
======================================================

@@

Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials. The use of genetic algorithms to find the parameters of a polynomial model is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.

@@

Introduction
------------

With notable exceptions, most machine learning techniques are based on linear models. This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework. However many application scenarios pose nonlinear problems that have no acceptable linear approximations.  
GIVE EXAMPLES AND EXPLAIN WHY LINEAR MODELS FAIL IN THOSE EXAMPLES.

Polynomials, one of the most studied subjects in mathematics, generalize linear models and are, perhaps, one of the simplest and most used nonlinear models. Applications of (nonlinear) polynomial models include colorimetric calibration [APPL1-2005], finding explicit formulae for turbulent pipe flows [APPL2-1999], computational linguistics [APPL8-2009] and more recently, analytical techniques for cultural heritage materials [APPL3-2010], liquid epoxy molding process [APPL4-2011], B-spline surface reconstruction [APPL5-2012], product design [APPL6-2012] or forecasting cyanotoxins presence in water reservoirs [APPL7-2013]. Besides the huge range of quite different areas, the work for each one of these polynomial models used, somewhere, a genetic algorithm.

Genetic algorithms (GA), are perhaps one of the most popular "buzzwords" of recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application. If a nonlinear model requires parameterization, GAs provide a simple and often effective approach to search for localy optimal parameters. Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to GAs come from John Holland and his book [BOOK-JOHNHOLLAND]. One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza, proposed the use of GAs to search the syntatic stucture of complex functions. This syntatic structure search is keen to the central ideas of deep learning, the subarea of machine learning actually producing the most advanced results.  
INSERT MORE REFERENCES OF RECENT GA RESEARCH AND DEEPLEARNING RESULTS

**END** STATE-OF-THE-ART

SUMMARIZE OUR PROBLEM STATEMENT, EXPLAIN WHY IT IS IMPORTANT, OUR RESOLUTION, EVALUATION STRATEGY AND RESULTS.

Polynomial Regression with Genetic Algorithms
---------------------------------------------

HOW DO WE DO THE
* polynomial encoding
* cost function
* genetic operators
* selection of other search parameters (_eg_ population size, maxiter, _etc_)

@@

Results
-------

@@

Conclusion
----------

References
----------

@@

[GAPOLY1-2011] http://dx.doi.org/10.4203/ccp.97.39 "Optimal Polynomial Regression Models by using a Genetic Algorithm"

[GAPOLY2-2008] http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.152.7415 "Optimal Sampling of Genetic Algorithms on Polynomial Regression"

[GAPOLY3-2011] http://www.eejournal.ktu.lt/index.php/elt/article/view/460 "Polynomial Curve Fitting with Varying Real Powers"

[GAPOLY4-2009] http://dx.doi.org/10.1016/j.eswa.2008.06.046 "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression"

[GAPOLY5-2006] http://dx.doi.org/10.1007/s00500-005-0008-8 "Genetic polynomial regression as input selection algorithm for non-linear identification"

@@

>>>>

# Learning Polynomial Regression with Genetic Algorithms

@@

Many applications require models that have no acceptable linear approximation and many nonlinear models are defined by polynomials. The use of genetic algorithms to find the parameters of a polynomial model is decades old but still poses challenges due to the complexity of the search and different definitions of "optimal" solution.
This work describes a general method based in genetic algorithms to find "empirical" polynomial regressions.

@@

## Introduction

With notable exceptions (_e.g._ $k$-means), most machine learning techniques are based on linear models. This assumption has many advantages including, for example, reduced computational complexity and strong theoretical framework. However nonlinearity in unavoidable in many application scenarions or, _e.g._ in systems with feedback loops so common in ecology, cybernetics, robotics and other areas.

Polynomials, one of the most studied subjects in mathematics, generalize linear models and are, perhaps, one of the simplest and most used nonlinear models. Applications of (nonlinear) polynomial models include colorimetric calibration ([APPL1-2005]), finding explicit formulae for turbulent pipe flows ([APPL2-1999]), computational linguistics ([APPL8-2009]) and more recently, analytical techniques for cultural heritage materials ([APPL3-2010]), liquid epoxy molding process ([APPL4-2011]), B-spline surface reconstruction ([APPL5-2012]), product design ([APPL6-2012]) or forecasting cyanotoxins presence in water reservoirs ([APPL7-2013]). Besides the huge range of quite different areas, the work for each one of these polynomial models used, somewhere, a genetic algorithm.

Genetic algorithms (GA), are perhaps one of the most popular "buzzwords" of recent decades but with good reason since they outline an optimization scheme easy to conceptualize and with very broad application. If a nonlinear model requires parameterization GAs provide a simple and often effective approach to search for localy optimal parameters. Research related to genetic algorithms abound and spans from the 1950's seminal work of Nils Aall Barricelli ([BARRICELLI-1962]) in the Institute for Advanced Study of Princeton to today's principal area of study of thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to GAs come from John Holland and his book ([BOOK-JOHNHOLLAND]).

One interesting "flavour" of genetic algorithms, termed "genetic programming" by John Koza ([KOZA1-1992]), proposed the use of GAs to search the syntatic stucture of complex functions. This syntatic structure search is keen to the central ideas of deep learning ([DEEPLEARNING1-2012] and [DEEPLEARNING2-2009]), the subarea of machine learning actually producing the most promising results. It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure ($y=\sum_i \beta_i x_i$) nonlinear (in particular polynomial) models pose an aditional "structure" search problem.

The idea of using GAs to find a polynomial regression is not new ([GAPOLY5-2006], [GAPOLY2-2008] and [GAPOLY4-2009]) but still generates original research ([GAPOLY1-2011] and [GAPOLY3-2011]). In line with that research this work describes a general method to find a polynomial regression of a given dataset. The optimal regression minimizes a cost function that accounts for both the _mean square error_ and a _regularization_ factor to avoid overfitting by penalizing polynomials that are "too complex".

A method that produces complex models "directly" from observed data has many uses. For example by a scientist to better understand the source of the data or by an autonomous agents adapting to the environment.  
MORE REASONS?

[OUR RESOLUTION]  
We represent a polynomial by a sum of monomials

$$
p\left( x_1, \ldots, x_n\right) = \sum_i \beta_i m_i
$$
where each monomial is $m_i = \prod_{j\in M_i} x_j^{\alpha_{ij}}$ the exponents are $\alpha_{ij}\in\mathbb{N}_0$ and the coeficients are $\beta_i \in \mathbb{R}$. For example 
$$
p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3
$$
has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\beta_1 = 2, \beta_2 = 1$ and $\beta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$.

This makes the problem of structure search very clear: except for the trivial cases, the number of possible monomials given $n$ variables and a maximum joint degree $d$ grows exponentialy with either $n$ or $d$. More importantly, the polynomial regression problem can be split into two subproblems:

1. for a given set of monomials $q_1, \ldots, q_n$, find regression coefficients $\beta_1,\ldots,\beta_n$ that minimize an adequate cost function;
2. find the fittest set of monomials;

Our line of work is to solve the first problem with linear regression and the second with GAs.

[EVALUATION STRATEGY]

[RESULTS]

## Polynomial Regression with Genetic Algorithms

HOW DO WE DO THE

* polynomial encoding
* cost function
* genetic operators
* selection of other search parameters (_eg_ population size, maxiter, _etc_)

@@

## Results

@@

## Conclusion

## References

**the references are defined "here" in the source but hidden in the resulting documents**

@@

[GAPOLY1-2011]: http://dx.doi.org/10.4203/ccp.97.39 "Optimal Polynomial Regression Models by using a Genetic Algorithm"

[GAPOLY2-2008]: http://citeseerx.ist.psu.edu/viewdoc/summary?doi=10.1.1.152.7415 "Optimal Sampling of Genetic Algorithms on Polynomial Regression"

[GAPOLY3-2011]: http://www.eejournal.ktu.lt/index.php/elt/article/view/460 "Polynomial Curve Fitting with Varying Real Powers"

[GAPOLY4-2009]: http://dx.doi.org/10.1016/j.eswa.2008.06.046 "A Novel hybrid genetic algorithm for kernel function and parameter optimization in support vector regression"

[GAPOLY5-2006]: http://dx.doi.org/10.1007/s00500-005-0008-8 "Genetic polynomial regression as input selection algorithm for non-linear identification"

@@

[DEEPLEARNING1-2012]: http://arxiv.org/abs/1206.5538 "Representation Learning: A Review and New Perspectives"

[DEEPLEARNING2-2009]: http://www.iro.umontreal.ca/~lisa/publications2/index.php/publications/show/239 "Learning Deep Architectures for AI"

[KOZA1-1992]: http://mitpress.mit.edu/books/genetic-programming "Genetic Programming: On the Programming of Computers by Means of Natural Selection, by John R. Koza, (1992)"

[BARRICELLI-1962]: http://dx.doi.org/10.1007%2FBF01556771 "Numerical testing of evolution theories : Part I Theoretical introduction and basic tests"
